Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-09-09 manuscript.tex
Draft second round of reveiew responses
round 2 review comments

\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\blauw}[1]{{#1}}

\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).} Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. \rood{With the exception of the simulation data reported in Section 3.1 where we draw comparisons against experimental data collected over a range of temperatures,} each sequence was simulated at its \rood{respective melting temperature dictated by the 3SPN.2 model} -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the nearest neighbor (NN) model predicted melting temperature,~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} and selecting the temperature at which the hybridized and dissociated populations were approximately equal. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined based on the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native Watson-Crick (WC) bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms}.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work, MSMs were constructed for each of the four DNA sequences \rood{at their respective 3SPN.2 melting temperatures} from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and defines the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.} We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box and ensure that we will observe asymptotically correct behavior at long times, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ as,}

    \begin{equation}\label{}
    \phi = \frac{f_\mathrm{hybridized}^\mathrm{sim}}{1 - f_\mathrm{hybridized}^\mathrm{sim}}
    \end{equation}

    \begin{equation}\label{}
	f_\mathrm{hybridized} = (1+\frac{1}{4\phi}) - \sqrt{(1+\frac{1}{4\phi})^2 - 1}
	\end{equation}

@@

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values in Fig.~\ref{fig:relaxation-comparison}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a \rood{universal (+4) K corrective temperature shift} to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected \rood{by these calibration corrections or finite-size fluctuation corrections} since we are only interested in the relative trends in the behaviors of the four sequences.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption{Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction at long times.} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

@@

We now proceeded to construct Markov state models (MSMs) from 1 ms of aggregated simulation trajectories for each of the four sequences at their respective melting temperatures to generate sequence-dependent kinetic models. MSMs define the long-lived metastable macrostates of the system, their equilibrium occupancies, and the equilibrium transition probabilities between them. As such, they are extremely valuable in providing both a quantitative predictive model and a physically comprehensible mechanistic understanding of the long-time dynamical evolution of the system between an ensemble of metastable macrostates. We present in Fig.~\ref{fig:allseq_table} the inferred MSMs for each of the four 10-base DNA sequences. Across all four sequences we identify a totality of seven metastable macrostates corresponding to the fully hybridized state (H) in which all native base pairings are intact, four shifted states in which the strands are translated out-of-register by two or four bases in the 5$^\prime$ (5S2, 5S4) or 3$^\prime$ (3S2, 3S4) direction, a frayed state (F4) -- unique to GC-core -- in which four terminal A:T base pairs are unbound, and the fully dissociated state (D). We present these seven macrostates in Fig.~\ref{fig:allseq_table}a along with schematic and cartoon renderings of representative microstates contained within each of these macrostates. \rood{The representative microstates emblematic of each macrostate were selected randomly from those possessing high ($>$99\%) macrostate membership probabilities. Since the microstate ensemble exhibits a range of conformations, however, it is useful to characterize the degree of structural heterogeneity to determine the degree to which these microstates are emblematic of the distribution. In \blauw{Fig.~S5} we present a projection of our macrostates into two physical order parameters -- the degree of 3$^\prime$ shift and degree of 5$^\prime$ shift -- to provide an interpretable embedding of the macrostates that exposes base pairing patterns and structural heterogeneity within the microstate ensemble. In all cases, we find these microstate distributions to be relatively narrowly focused within the low-free energy core of each macrostate such that a single archetypal microstate is indeed a good representative for the ensemble and an accurate representation of the base pairing pattern of the macrostate.} In Fig.~\ref{fig:allseq_table}b we present the occupancy probabilities of each state at thermodynamic equilibrium. By virtue of the fact that each sequence is simulated at its corresponding melting temperature (T$_\mathrm{m}$), the probability of the dissociated state (D) is, by construction, approximately equal to the sum of the probabilities over the remaining six states (H, 5S2, 3S2, 5S4, 3S4, F\rood{4}). In Fig.~\ref{fig:allseq_table}c we present a visualization of the macrostate MSMs for each of the four sequences showing the connectivity between the identified macrostates. The macrostates are represented as orange circles in proportion to their equilibrium probabilities and the grey arrows indicate the probability of hopping from one macrostate to another under one time step of the kinetic model. The fluxes between the macrostates provide a wealth of high-level, interpretable information on the sequence-dependent metastable states and hybridization/dehybridization pathways. Immediately, we identify that the AT-all sequence possesses a rich and complex dynamical landscape comprising six metastable states whereas at the other end of the spectrum GC-mix exhibits far simpler two-state ``all-or-nothing'' behavior. In Fig.~\ref{fig:allseq_table}d we present the so-called implied time scales of each MSM. These time scales correspond to the relaxation times of the DNA dimer among its constituent metastable macrostates. The leading implied time scale for each system corresponds to the characteristic time scale for hybridization/dehybridization. Since each system is simulated at the same concentration and at its respective melting temperature, it is not surprising that the leading time scale is approximately equal for all four systems and corresponds to the characteristic time scale for hybridization/dehybridization. The spectrum of higher order time scales corresponds to increasingly quicker relaxations between the metastable macrostates within the kinetic model and resolve the interesting sequence-dependent differences in the hybridization/dehybridization kinetic pathways. The total number of implied time scales is typically one fewer than the number of metastable macrostates and the existence of large implied time scales is indicative of slowly relaxing kinetic processes. The dense spectrum of slow implied time scales for AT-all is indicative of its relatively complex kinetic landscape whereas that for the two-state GC-mix comprises only a single time scale corresponding to hybridization/dehybridization. We now proceed to analyze in detail the sequence-dependent thermodynamics, kinetic, and mechanisms exposed by the four MSMs.

>>>>

%\newcommand*{\rood}[1]{{\color{red}{#1}}}
\newcommand*{\rood}[1]{{#1}}
%\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\blauw}[1]{{#1}}
\newcommand*{\roodr}[1]{{\color{red}{#1}}}

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
%\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

All calculations were performed using the LAMMPS simulation package (\url{http://lammps.sandia.gov}) in accordance with best practices for the 3SPN.2 model. \citep{Plimpton1997FastDynamics} A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).} Solvent effects were modeled implicitly by employing Langevin dynamics \citep{DunwegB.Paul1991BrownianNumbers., Bussi2007AccurateDynamics} with an experimentally motivated per-site friction coefficient of 9.94 $\times$ 10$^{-11}$ m$^2$/s. \citep{Nkodo2001DiffusionElectrophoresis, Hinckley2013AnHybridization} We specified a 240 mM implicit salt concentration and treated electrostatic interactions using the Debye-H{\"u}ckel with a 5 nm cutoff radius. \citep{Debye1923ZurElektrolyte} Simulations were performed in the NVT ensemble employing a Langevin thermostat. \citep{Schneider1978Molecular-dynamicsTransitions}. \rood{With the exception of the simulation data reported in Section 3.1 where we draw comparisons against experimental data collected over a range of temperatures,} each sequence was simulated at its \rood{respective melting temperature at 7mM concentration as dictated by the 3SPN.2 model} -- AT-all: 309 K, GC-end: 317 K, GC-core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states. Melting temperatures for each sequence were determined empirically by comparing the ratio of hybridized to dissociated populations over a 10 K temperature ramp centered on the nearest neighbor (NN) model predicted melting temperature,~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} and selecting the temperature at which the hybridized and dissociated populations were approximately equal. The Langevin equations of motion were integrated using the scheme of Bussi and Parrinello \citep{Bussi2007AccurateDynamics} with a 20 fs integration time step. We performed 40 independent simulations for each of the four sequences with half of the runs initialized from the hybridized state and half from the dissociated state. The initial hybridized state was defined based on the crystal structure coordinates of Arnott et al.~\citep{Arnott1976CRCBiology} The dissociated state was generated from the hybridized state by displacing one strand away from the other by 1 nm in each of the $x$, $y$, and $z$ directions. The magnitude of this displacement did not affect the results so long as all native Watson-Crick (WC) bonds were completely broken. Initial bead velocities were assigned from a Maxwell-Boltzmann distribution at the temperature of interest. Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.

@@

Markov state models (MSMs) are a powerful approach to infer long-time kinetic models from short molecular simulation trajectories \citep{Sengupta2019AutomatedSelf-assembly, Pande2010EverythingAsk, Chodera2014MarkovDynamics, Husic2018MarkovScience, Wehmeyer2019IntroductionSoftware} that we employ in this work to construct high-resolution sequence-dependent kinetic models of DNA hybridization and dissociation. In brief, a MSM extracts from simulation trajectories an ensemble of long-lived metastable macrostates, their equilibrium occupancy populations, and the equilibrium probability fluxes between them. In this manner, they provide an interpretable and predictive model of the system thermodynamics and kinetics. MSMs have recently been implemented to study the hybridization mechanism of one particular 14-mer DNA oligomer, but determining the sequence-dependent kinetics and mechanisms was not the focus of this study.\citep{Jin2019} An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms}.~\citep{Xiao2019} Kinetic models were constructed not from the dynamical trajectories of the molecular model, but by estimating rate constants between local minima using a transition state theory approximation. A recent application of MSMs to the long-time dynamics of short RNA oligonucleotides revealed stacking time scales to be highly sequence dependent.~\citep{Pinamonti2017} In this work,  MSMs were constructed for each of the four DNA sequences \roodr{at their respective 3SPN.2 melting temperatures at 7 mM concentration} from the 40$\times$25 $\mu$s simulation trajectories following a four-step protocol detailed in Ref.~\citep{Sidky2019High-ResolutionVAMPnets}: (i) trajectory featurization, (ii) dimensionality reduction, (iii) microstate clustering and microstate transition matrix inference, (iv) macrostate clustering and macrostate transition matrix inference. Calculations were performed using the PyEMMA software package. \citep{Scherer2015PyEMMAModels}

@@

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of \roodr{2.0} nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and defines the dissociated state. \roodr{We verified that our results were robust to choices of this cutoff over the range 1.3-2.5 nm. as measured by a <7\% average deviation in rates over that range.} We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$.

%\rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box and ensure that we will observe asymptotically correct behavior at long times, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ as,}

%    \begin{equation}\label{}
%    \phi = \frac{f_\mathrm{hybridized}^\mathrm{sim}}{1 - f_\mathrm{hybridized}^\mathrm{sim}}
%    \end{equation}

%    \begin{equation}\label{}
%	f_\mathrm{hybridized} = (1+\frac{1}{4\phi}) - \sqrt{(1+\frac{1}{4\phi})^2 - 1}
%	\end{equation}

@@

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values in Fig.~\ref{fig:relaxation-comparison}. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a \rood{universal (+4) K corrective temperature shift} to our computational results. \roodr{We note that we would expect the model to slightly over-predict melting temperature due to the 3.5$\times$ higher concentration relative to experiment, however the systematic shift shows that relative sequence-dependent trends are well maintained. Furthermore,} We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but acknowledge that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not \roodr{corrected by these calibration corrections} since we are only interested in the relative trends in the behaviors of the four sequences.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption{Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

@@

We now proceeded to construct Markov state models (MSMs) from 1 ms of aggregated simulation trajectories for each of the four sequences at their respective melting temperatures to generate sequence-dependent kinetic models. MSMs define the long-lived metastable macrostates of the system, their equilibrium occupancies, and the equilibrium transition probabilities between them. As such, they are extremely valuable in providing both a quantitative predictive model and a physically comprehensible mechanistic understanding of the long-time dynamical evolution of the system between an ensemble of metastable macrostates. We present in Fig.~\ref{fig:allseq_table} the inferred MSMs for each of the four 10-base DNA sequences. Across all four sequences we identify a totality of seven metastable macrostates corresponding to the fully hybridized state (H) in which all native base pairings are intact, four shifted states in which the strands are translated out-of-register by two or four bases in the 5$^\prime$ (5S2, 5S4) or 3$^\prime$ (3S2, 3S4) direction, a frayed state (F4) -- unique to GC-core -- in which four terminal A:T base pairs are unbound, and the fully dissociated state (D). We present these seven macrostates in Fig.~\ref{fig:allseq_table}a along with schematic and cartoon renderings of representative microstates contained within each of these macrostates. \rood{The representative microstates emblematic of each macrostate were selected randomly from those possessing high ($>$99\%) macrostate membership probabilities. Since the microstate ensemble exhibits a range of conformations, however, it is useful to characterize the degree of structural heterogeneity to determine the degree to which these microstates are emblematic of the distribution. In \blauw{Fig.~S5} we present a projection of our macrostates into two physical order parameters -- the degree of 3$^\prime$ shift and degree of 5$^\prime$ shift -- to provide an interpretable embedding of the macrostates that exposes base pairing patterns and structural heterogeneity within the microstate ensemble. In all cases, we find these microstate distributions to be relatively narrowly focused within the low-free energy core of each macrostate such that a single archetypal microstate is indeed a good representative for the ensemble and an accurate representation of the base pairing pattern of the macrostate.} In Fig.~\ref{fig:allseq_table}b we present the occupancy probabilities of each state at thermodynamic equilibrium. By virtue of the fact that each sequence is simulated at its corresponding melting temperature (T$_\mathrm{m}$), the probability of the dissociated state (D) is, by construction, approximately equal to the sum of the probabilities over the remaining six states (H, 5S2, 3S2, 5S4, 3S4, F\rood{4}). In Fig.~\ref{fig:allseq_table}c we present a visualization of the macrostate MSMs for each of the four sequences showing the connectivity between the identified macrostates. The macrostates are represented as orange circles in proportion to their equilibrium probabilities and the grey arrows indicate the probability of hopping from one macrostate to another under one time step of the kinetic model. The fluxes between the macrostates provide a wealth of high-level, interpretable information on the sequence-dependent metastable states and hybridization/dehybridization pathways. Immediately, we identify that the AT-all sequence possesses a rich and complex dynamical landscape comprising six metastable states whereas at the other end of the spectrum GC-mix exhibits far simpler two-state ``all-or-nothing'' behavior. In Fig.~\ref{fig:allseq_table}d we present the so-called implied time scales of each MSM. These time scales correspond to the relaxation times of the DNA dimer among its constituent metastable macrostates. The leading implied time scale for each system corresponds to the characteristic time scale for hybridization/dehybridization. Since each system is simulated at the \roodr{same 7 mM concentration} and at its respective melting temperature, it is not surprising that the leading time scale is approximately equal for all four systems and corresponds to the characteristic time scale for hybridization/dehybridization. The spectrum of higher order time scales corresponds to increasingly quicker relaxations between the metastable macrostates within the kinetic model and resolve the interesting sequence-dependent differences in the hybridization/dehybridization kinetic pathways. The total number of implied time scales is typically one fewer than the number of metastable macrostates and the existence of large implied time scales is indicative of slowly relaxing kinetic processes. The dense spectrum of slow implied time scales for AT-all is indicative of its relatively complex kinetic landscape whereas that for the two-state GC-mix comprises only a single time scale corresponding to hybridization/dehybridization. We now proceed to analyze in detail the sequence-dependent thermodynamics, kinetic, and mechanisms exposed by the four MSMs.
